Evaluating Modern vs. Older Buildings: Detailed Analysis on the Apartment Building Evaluation 2023

### 2024-01-24 Output/Paper/BuildingScoreAnalyze-2023.qmd
Update on TA's Suggestion

To fully understand the relationship between the year of construction and the evaluation score, I have divided this paper into data, results, discussion, and conclusion sections. All data were obtained from the OpenDataToronto package [@citeopendata]. In the data section, I will select relevant parts of my cleaned data to explain what each column represents. The results section will present all the graphs useful in analyzing the relationship. In the discussion section, I will interpret these graphs and offer my conclusions based on the provided data. Finally, the conclusion will summarize the entire paper and my main points.

@@

```{r, fig.height=3, fig.width=5, fig.align='center'}
#| message: false
#| echo: false
#| label: fig-fig1
#| fig-cap: Comparing Year with Respect of Score

@@

Since we cannot draw definitive conclusions from the graph above, creating a graph that calculates the mean score for each decade can provide a much clearer understanding of the overall pattern or relationship between year and score. The following graph, which displays the relationship between decade and average score per decade (see @fig-fig2), reveals an obvious upward trend. Given the limited number of data points before 1900-possibly due to many older buildings being replaced with modern constructions-it is better to focus on data from post-1900.

```{r, fig.height=2, fig.width=4, fig.align='center', fig.cap=""}
#| message: false
#| echo: false
#| label: fig-fig2
#| fig-cap: Comparing decade with average scores

## create a graph of decade vs. average score with help of Chatgpt
ggplot(bu_eval, aes(x = Decade, y = Average_Score)) +
  geom_point() +
  labs(x = "decade", y = "average scores", title = "Plot of decades on average scores") +
  theme_minimal()
```

@@

```{r, fig.height=3, fig.width=5, fig.align='center'}
#| message: false
#| echo: false
#| label: fig-fig3
#| fig-cap: Comparing Category with Percentage of Achieveing Full Mark

@@

Additionally, I was keen to explore the relationship of scores based on unit or storey numbers. After plotting graphs of scores against unit numbers (see @fig-fig4) and storey numbers (see @fig-fig5), a similar pattern emerges in both graphs. For buildings with a lower count of units (under 50) and storeys (under 10), scores vary widely, ranging from below 25 to 100. Conversely, buildings with a higher count of units (over 200) and storeys (above 20) tend to have scores consistently above 75.

```{r, fig.height=3, fig.width=5, fig.align='center'}
#| message: false
#| echo: false
#| label: fig-fig4
#| fig-cap: Comparing Score with Unit Number

## create a graph of unit number vs. score with help of Chatgpt
ggplot(bu_eval, aes(x = units, y = score)) +
  geom_point() +
  labs(x = "Unit number", y = "Score", title = "Score vs. Unit number") +
  theme_minimal()
```

```{r, fig.height=3, fig.width=5, fig.align='center', fig.cap=""}
#| message: false
#| echo: false
#| label: fig-fig5
#| fig-cap: Comparing Score with Storey Number

## create a graph of storeys number vs. score with help of Chatgpt
ggplot(bu_eval, aes(x = storeys, y = score)) +
  geom_point() +
  labs(x = "Storey number", y = "Score", title = "Score vs. storey number") +
  theme_minimal()
```

@@

@@

[^2]:**Note: After Peer Review, a friend commented my paper and said that outliers are serious issue and suggest that I take care of it. I redid my codes and removed 1% each side of my data on column score and I found that my conclusion stayed the same. The graph is still having a trend I talked about above. Thanks to @ zxc0707.

@@

**Reference From ref.bib:

>>>>

To fully understand the relationship between the year of construction and the evaluation score, I have divided this paper into data, results, discussion, and conclusion sections. All data were obtained from the OpenDataToronto package [@citeopendata]. In the data section, I will select relevant parts of my cleaned data to explain what each column represents. The results section will present all the graphs useful in analyzing the relationship. In the discussion section, I will interpret these graphs and offer my conclusions and suggestions based on the provided data which to give renters some valuable advise of which building is better to pick, old or new, high or low. Finally, the conclusion will summarize the entire paper and my main points.

@@

```{r, fig.height=3, fig.width=5, fig.align='center'}
#| message: false
#| echo: false
#| label: fig-fig1
#| fig-cap: How Age can Effect a Building's Score

@@

```{r}
#| message: false
#| echo: false
# Only show decade after 1900
bu_eval_filtered <- bu_eval %>% filter(Decade >= 1900)
# Create a model 
model <- lm(Average_Score ~ Decade, data = bu_eval_filtered)
#Calculate coefficients
coefficients <- coef(model)
# put coefficients to variables
intercept_pink <- coefficients[1]
slope_pink <- coefficients[2]

slope_pink_mod <- paste("k =", round(slope_pink, digits = 3))
equation_pink <- paste("y = ", round(slope_pink, digits = 2), "x + ", round(intercept_pink, digits = 2))

```

Since we cannot draw definitive conclusions from the graph above, creating a graph that calculates the mean score for each decade can provide a much clearer understanding of the overall pattern or relationship between year and score. The following graph, which displays the relationship between decade and average score per decade (see @fig-fig2), reveals an obvious upward trend. Given the limited number of data points before 1900-possibly due to many older buildings being replaced with modern constructions-it is better to focus on data from post-1900. I added a best fit line which will show the big trend of the point graph. I also calculated the y-intersection and slope for the line which is `r equation_pink`. I think in this case the y-intersection is actually useless while slope is more useful. We can see from the slope `r slope_pink_mod` that in general for every 10 years of age for a building, it is going to lose 1.3 points which seems like not so much.

```{r, fig.height=2, fig.width=4, fig.align='center'}
#| message: false
#| echo: false
#| label: fig-fig2
#| fig-cap: Average Score for each Decade which a building is built on

## create a graph of decade vs. average score with help of Chatgpt
ggplot(bu_eval_filtered, aes(x = Decade, y = Average_Score)) +
  geom_point() +
  geom_smooth(method = "lm", color = "pink") +
  labs(x = "decade", y = "average scores", title = "Decades on average scores") +
  theme_minimal()
```

@@

```{r, fig.height=3, fig.width=5, fig.align='center'}
#| message: false
#| echo: false
#| label: fig-fig3
#| fig-cap: Comparing Each Category with the Percentage of a Building in its Decade of Achieveing Full Mark

@@

# Create a model 
model_blue <- lm(score ~ units, data = bu_eval)
#Calculate coefficients
coefficients <- coef(model_blue)
# put coefficients to variables
intercept_blue <- coefficients[1]
slope_blue <- coefficients[2]

slope_blue_mod <- paste("k =", round(slope_blue, digits = 3))
equation_blue <- paste("y = ", round(slope_blue, digits = 2), "x + ", round(intercept_blue, digits = 2))
```

# Create a model 
model_green <- lm(score ~ storeys, data = bu_eval)
#Calculate coefficients
coefficients <- coef(model_green)
# put coefficients to variables
intercept_green <- coefficients[1]
slope_green <- coefficients[2]

slope_green_mod <- paste("k =", round(slope_green, digits = 3))
equation_green <- paste("y = ", round(slope_green, digits = 2), "x + ", round(intercept_green, digits = 2))
```

Additionally, I was keen to explore the relationship of scores based on unit or storey numbers. After plotting graphs of scores against unit numbers (see @fig-fig4) and storey numbers (see @fig-fig5), a similar pattern emerges in both graphs. For buildings with a lower count of units (under 50) and storeys (under 10), scores vary widely, ranging from below 25 to 100. Conversely, buildings with a higher count of units (over 200) and storeys (above 20) tend to have scores consistently above 75. By adding a best fit line `r equation_blue` for unit number and `r equation_green`, we can see that it has a slope of `r slope_blue_mod` which we can say that it is almost a horizontal line. Also for the relationship between storey and score, even though that the slope is `r slope_green_mod` which is not necessary horiziontal line, but it still does not provide us with a relationship.

```{r, fig.height=3, fig.width=5, fig.align='center'}
#| message: false
#| echo: false
#| label: fig-fig4
#| fig-cap: How number of Units Effect a Building's Score

## create a graph of unit number vs. score with help of Chatgpt
ggplot(bu_eval, aes(x = units, y = score)) +
  geom_point() +
  geom_smooth(method = "lm", color = "blue") +  # Adds a best fit line
  labs(x = "Unit number", y = "Score", title = "Score vs. Unit number") +
  theme_minimal()
```

```{r, fig.height=3, fig.width=5, fig.align='center', fig.cap=""}
#| message: false
#| echo: false
#| label: fig-fig5
#| fig-cap: How Number of Storeys will Effect a Building's Score

## create a graph of storeys number vs. score with help of Chatgpt
ggplot(bu_eval, aes(x = storeys, y = score)) +
  geom_point() +
  geom_smooth(method = "lm", color = "green") +
  labs(x = "Storey number", y = "Score", title = "Score vs. storey number") +
  theme_minimal()
```

@@

[^2]: \*\*Note: After Peer Review, a friend commented my paper and said that outliers are serious issue and suggest that I take care of it. I redid my codes and removed 1% each side of my data on column score and I found that my conclusion stayed the same. The graph is still having a trend I talked about above. Thanks to \@ zxc0707.

@@

@@

\*\*Reference From ref.bib:
